---
title: "Bayesian Trend Filtering"
author: "Edward A. Roualdes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: refs.bib
vignette: >
  %\VignetteIndexEntry{btf}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Introduction

Bayesian trend filtering estiamtes the function $f = f(x_1), \ldots, f(x_n)$ from data $y_i, i=1, \ldots, n$, hypothesized to be generated from the model $$y_i = f(x_i) + \epsilon_i$$ by fitting a fully Bayesian hierarchical model that is equivalent to the minimization problem $$argmin_f ||y - f||_2^2 + \lambda||D^k f||_1$$ as detailed by R.J. Tibshirani [-@Tibshirani:2014].  The penalty matrix $D^k$ estimates $k$th derivatives of $f$ at various inputs $x_i$.  This is the Bayesian analogue to the function [`genlasso::trendfilter`](http://cran.r-project.org/web/packages/genlasso/index.html).

The main function of this package, `btf::btf`, assumes the inputs $x_i, i=1, \ldots, n$, are equally spaced and in ascending order.  The user needs to specify the order of the derivatives $k$.  I recommend $k<=3$ since calculations with $D^k$ can otherwise cause numerical instability.  Note that unlike `genlasso::trendfilter` Bayesian trend filtering will not set any elements of the penalty vector $D^k f$ exactly equal to zero and thus this method should be used as a smoother.

## Fitting btf

We'll use the time-series observations `datasets::nhtemp`.  First, fit a model by specifying the observations $y$ and the order of the derivative to use.

```{r}
library(btf)
fit <- btf(nhtemp, k=3)
```

The variable `fit` is nothing special, but I recommend retrieving the posterior samples with the functions `btf::getPost`.  You can specify which parameter you want to investigate, by default we assume you are interested in the estimate of $f$.  Output from this function is a `coda::mcmc` object, suitable for any of the methods specified within the [`coda`](http://cran.r-project.org/web/packages/coda/index.html) package.

```{r, fig.show='hold', fig.width=7, fig.height=5}
plot(getPost(fit, 's2'))
```

Alternatively, you can ask for a summary of the parameters requested.  The function `btf::getPostEst` will give the output from `coda::summary.mcmc` applied to the parameters specified, or $f$ is no paramters are specified.  You can also specify a function be applied to each of the paramters requested.

```{r}
getPostEst(fit, 's2')
getPostEst(fit, 's2', est=mean)
```

A simple plot of the estimate of $f$ and highest posterior density intervals is provided by the S3 method `btf::plot.btf`

```{r, fig.show='hold', fig.width=5, fig.height=5}
plot(fit)

```

## References
